%---------------------------------------------------------------------------------
\chapter{Literature Survey}
\label{chap:Literature-Survey}
%---------------------------------------------------------------------------------

As a critical modules for many speech processing applications, VAD has received considerable research interests in the last couple of decades. 
It started as early as in the 1970s, when VAD was often referred to as speech endpoint detection or word boundary detection problem. 
Back in that time, VAD algorithms often deal with only little or no noise corruption in speech coding applications, 
and with separate recording utterances in speech recognition systems. 
Up to recently, advances in various speech applications require the detection of human speech in a continuous real-time fashion, 
and is often corrupted by a wide variety classes of noise. Algorithms for VAD had grown accordingly over the years. 
Unfortunately, given the vast variety of VAD techniques proposed in the literature, no work has been found to sufficiently study and compare the state-of-the-art techniques. 
It is part of this research's contributions to provide an extensive and up-to-date literature survey for VAD.

The core of any VAD is proposed to be consisted of two parts: a `feature extraction' and a `speech/ non-speech decision mechanism.' 
The first part extracts from the given speech signal the parameters that can represent the discriminative characteristics of speech comparing to noise. 
Using these parameters, the second part makes the final speech/non-speech decision, based on a set of decision rules.

The rest of this chapter reviews these two processes in more details. 
It first introduces different features proposed in literature, divided into categories; and then, the approaches to speech/non-speech decision mechanism are studied.

%TODO: briefing section

\input{features}
\input{decision_rules}
